\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Submitted to Ecosphere. Manuscript type: Article}}

### 2023-10-10 manuscript/manuscriptRW.Rmd
more updates in the revision before sending to co-authors

Most of the models overestimate regeneration levels, which is only compensated in some models by high simulated mortality rates in the early stages of individual trees dynamics. Simulated species diversity of regeneration matches the observed ranges. Models simulating higher species diversity at the stand level do not present higher regeneration diversity. The effect of light availability on regeneration levels is captured better than the effect of temperature and soil moisture, but patterns are not consistent across models. Increasing complexity in the tree regeneration modules of the models is not related to higher accuracy of simulated tree regeneration. Furthermore, individual model design is more important than scale (stand, landscape, global) and approach (empirical, process-based) for accurately capturing tree regeneration.

@@

Thus, a focus on the modeling of tree regeneration processes is sorely needed and overdue [@price2001; @walck2011] if we are to make reliable projections of future forest dynamics, i.e., when the models need to be operated in extrapolation mode, as well as from a fundamental ecological point of view for increasing systems understanding. In the present study, for the first time, a large number of forest models commonly used to assess forest dynamics under climate change are evaluated against a continental-scale, multi-species harmonized dataset on tree regeneration [@käber2023]. By tree regeneration, we refer to the passing of trees across a specific diameter threshold ("ingrowth"). We included models that are based on a range of "philosophies", from models operating at the stand to the global scale as well as the range of models from empirically derived to "process-based" [e.g., @bugmann1996; @fabrika2012; @lexer2001; @reyer2014]. Due to the large variability in tree regeneration patterns in nature and the large number of factors driving and constraining this process --- including some that are not incorporated explicitly in most models, such as deer browsing --- we do not aim for a detailed statistical evaluation of each model. Instead, we aim to evaluate the general regeneration patterns and magnitudes simulated by the models, and benchmark the simulated regeneration niche of multiple species against empirical data along a wide environmental gradient of temperature, moisture, and light availability [@grubb1977].

More specifically, we aim to answer the following questions: (1) Are models of forest dynamics capturing accurately tree regeneration levels, initial tree species diversity and mortality in the regeneration? (2) Do model traits explain differences in model performance? (3) How well do the models capture total regeneration and individual species regeneration niches across environmental gradients of light availability, temperature, and soil moisture? We evaluate the performance of regeneration models by analyzing simulated data alongside observed data. Specifically, we examine tree species diversity, mortality rates in tree regeneration, and regeneration levels. Then, we link the model's traits and performance to define regeneration and we assess the regeneration gradients on regeneration niches. Finally, we provide research recommendations for next steps towards improving tree regeneration modules.

@@

In the models used here, the overall complexity in the regeneration modules varies considerably. Following @bugmann2022, we can classify seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and TreeMig) as having rather high complexity in their regeneration modules (mean regeneration formulation complexity across all processes \>0, Table \@ref(tab:tableModels)). Two models, ForClim and ForCEEPS, were used here with two alternative variants of regeneration. ForClim variant 1 [@bugmann1996] is based on a recruitment module that adheres closely to the concept introduced by @botkin1972 which is an environmental filter considering species specific thresholds of light availability and climatic variables. Whereas ForClim variant 11 is adopting a slightly more complex approach where individual species properties and their relation to the environment are represented gradually along with the relative suitability of different species (i.e., competition among tree species) [@huber2020]. These two model variants allow us to evaluate a more process-based and complex module (variant 11) against a simple module (variant 1), while the rest of the model structure is identical. Similarly, the two ForCEEPS variants allow us to isolate the importance of the canopy feedback (i.e., simulated actual composition and relative abundance of species in the plot) via seed trees for the quantity and quality (e.g., diversity and composition) of simulated regeneration, as one variant includes this feedback whereas the other does not.

@@

Regeneration data covering a wide range of environmental conditions are hard to obtain, and this is one of the reasons why most models of forest dynamics have never been confronted with a dataset covering such gradients over a large number of sites, to evaluate how well regeneration is captured. The observations used here are derived from a novel network of sites in forest reserves that represent the range of environmental gradients in temperature and precipitation in Central Europe as compiled in the framework of the EuFoRIa network [@euforia2019;TODO add @kaber2023] (Figure \@ref(fig:map)). The environmental coverage and the variety of forests represented by this data is unprecedented within Europe. These forests have been unmanaged for at least 10 years prior to their designation as reserves, and most of them provide time series of natural forest dynamics over multiple decades and up to ca. 80 years. The census periods range from 3 to 37 years, with an average of 14 years. The data provide information at the tree level, thus allowing for the sequential comparison of processes such as individual-tree regeneration and death between the measurements.

We selected 200 sites from this network as the benchmarking dataset for the simulation to be representative of the environmental variation contained in the data. This was achieved by applying k-means clustering to define 200 clusters of plots from the original set of 869 plots along the environmental dimensions of temperature, climatic water balance, soil quality, slope and aspect. Each of these sites featured at least two consecutive measurements. Regeneration minimum size thresholds for these sites differed between stem diameters of 0 and 10 cm. For our study, we defined two datasets where one included 165 sites with a diameter threshold of 7 cm or lower, and the other included another 35 sites with diameter thresholds from 10 cm. Therefore we had in total 200 sites with observations of newly established trees above 10 cm and 165 sites above 7cm. In the results section any result from observed data at 7 cm threshold is originated from the 165 sites.

Plot size ranged from 0.02 ha to 5.52 ha, and the data were further processed and aggregated following @käber2023 to provide regeneration rates per ha and per decade. The observations used in this study featured 30,900 newly established trees. Regeneration rates per site, sample and decade ranged from zero to 1246 trees, with a mean of 56 trees. Adult species composition was also available for each plot. For more details on this unique dataset and the detailed regeneration information, cf. @käber2023.

@@

The overarching goal of the simulation experiments was to assess tree regeneration as it arises from empirical data against its representation in a wide range of models of forest dynamics. We define tree regeneration as the passing of a breast height diameter threshold of 7 and 10 cm (synonym: ingrowth). To this end, each modeling group was provided with a detailed protocol [@díaz-yáñez2022] with instructions how to perform the simulations, input variables on climate and soil conditions, and the list of expected output variables. Neither were further site information (except for the data specified below) nor any data on tree regeneration or forest stand features provided prior to the simulation.

@@

The simulations were run in the absence of natural disturbances. Only the model LPJ-GUESS had to include a background disturbance to increase chances of shade-intolerant species to establish. The simulations were set up to sample species-specific regeneration rates per decade and per ha in the equilibrium state of the model, typically entailing a "spin-up" run (as we did not provide any forest data). The modeling teams decided on the simulated area and how they derived these samples. The exact length of the simulation was also decided by the modeling teams (Table \@ref(tab:tableModels)). Further details on how each modeling team prepared the simulations and the outputs are available in Supplementary Material 1.

The simulations were run in the absence of management to a simulated equilibrium ("Potential Natural Vegetation") with the current climate, as the ultimate goal was to evaluate tree regeneration under comparable and near-equilibrium conditions. This entails the assumption that (1) the observations from the forest reserves reflect no traces of forest management, and (2) there is an equilibrium between forest dynamics and climate. While the former might be starting to be visible in many of the EuFoRIA reserves, the latter may be more debatable. However, in the absence of detailed data on the history of each plot in the EuFoRIa network, some broad assumptions had to be made. Both the width of the regeneration niche (i.e., in environmental space) as well as the intensity of the regeneration process (i.e., the number of ingrowth trees per area and per unit of time) were of interest.

@@

Two models did not provide results from all the simulated samples or sites to avoid unrealistic results, as follows. In the model 4C, a threshold of a maximum basal area of 90 $m^2 ha^{-1}$ was used to avoid unrealistic stand basal area data, and therefore not all the samples and sites were reported. The reason for this is that 4C is not suitable for long-term simulations without management, due to misrepresentations in density-dependent mortality processes in long-term simulations and assumptions of tree geometry that lead to unrealistic single tree dimensions of very old individuals. LPJ-GUESS had 2% of the sample outputs without tree regeneration, and these were considered as zero stand basal area and zero regeneration for all the tree species simulated; one site produced grassland rather than a forest and was not included in the results.

@@

The simulation results were analyzed regarding (1) regeneration levels (i.e., ingrowth number per unit time and space), (2) regeneration species diversity, (3) regeneration mortality, (4) the relationship between model performance and model traits, and (5) ingrowth gradients along the regeneration niches. We evaluated species diversity across the models and in relation to the observed data by calculating the Shannon index $H$ based on the relative proportion of the species in terms of basal area. It was calculated for the regeneration ($H_{R_{n}}$) (Equation \ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$) (Equation \ref{eqpiShannon}). The higher the value of the index, the higher the species diversity at a particular site and sample.

@@

Mortality in tree regeneration was assessed based on the ratio of regeneration between the 7 and 10 cm diameter thresholds. We used the Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference to estimate whether the ratio of regeneration between the 7 and 10 cm diameter thresholds were above or below the expected theoretical rate. The Reineke self-thinning rule is usually calculated for even-aged, single species stands and assumes a fixed relationship between the number of stems and the quadratic mean diameter in fully stocked pure stands. The value used in our comparisons was 1.77 (i.e., we expect stem numbers at 7 cm to be 77% higher than at 10 cm), calculated using Equation \ref{eqReineke}.

```{=tex}
\begin{equation}
Reineke = \frac{N7}{N10}^{-1.605} \label{eqReineke}
\end{equation}
```
We assessed model performance in relation to model traits focusing on (1) model complexity as defined by @bugmann2022, (2) model type (empirical or process based), (3) the presence or absence of a canopy feedback for regeneration, and (4) the scale of application of the model (stand, landscape or global). We tested for significant differences using two-sided t-tests by considering each of these model features regarding the observed and simulated regeneration levels and species diversity, respectively.

Total regeneration and the regeneration niches of the individual species were evaluated across the environmental gradients of light availability, temperature, and soil moisture. Stand basal area (the basal area of all the trees in each sample including tree regeneration) was used as a proxy for light availability at the forest floor, the annual degree-day sum [@allen1976; @fischlin1995] as a proxy for growing season warmth, and the climatic water balance as a proxy for soil moisture [@speich2019]. Regeneration values were calculated as the mean across the 200 samples per site. The observed data were modeled using a Generalized Additive Model [@wood2011] with a negative binomial distribution and restricted maximum likelihood to better understand the relationship between the environmental gradients and the levels of observed regeneration, relative to the simulation results per model.

In order to analyze the regeneration niches across the climatic gradients, we focused on five common tree species or genera: *Abies alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and *Quercus* spp. For these species, we calculated the share in the regeneration basal area per site as the mean across the available samples per site:

```{=tex}
\begin{equation}
\bar{R}_{BA_{share_{t,i}}} = \frac{\bar{r}_{BA_{t,i}}}{\sum_{i = 1}^{s}{\bar{r}_{{BA_{t,i}}}}} \label{eq:BAshare}
\end{equation}
```
where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$ is the mean basal area in the regeneration (subsequently referred to as "regeneration basal area") of species $i$ at site $t$ across the available samples at that site. Furthermore, regarding the share in the regeneration basal area per species ($\bar{R}_{BA_{share_{i}}}$), we categorized this as zero when the regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) was zero, but also when both the total regeneration basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$) and the regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) equaled zero.

@@

Simulated regeneration levels varied strongly across the 15 models and typically did not match the levels found in the forest reserve data (Figure \@ref(fig:over)). Regeneration was overestimated in most models for both the 7 and 10 cm diameter thresholds, with the exception of the empirical stand model SIBYLA, the landscape model Landis-II and the global model aDGVM2, which estimated regeneration levels at the lower end of the plausibility interval of the observed data (Figure \@ref(fig:over)). The models with the largest overestimation were the stand model PICUS and the landscape model TreeMig. For most models, the variability of simulated regeneration levels across the 200 sites (visible from the interquartile range in the box plots of Figure \@ref(fig:over)) was similar to or smaller than observed, with the notable exception of PICUS, where simulated regeneration variability was much larger.

Both the observations and the simulated data had no regeneration in some samples and at some sites (for details, cf. Table S1). The observed data had 4% of the samples with no regeneration. Only three models had a larger proportion of no regeneration (4C, Landis-II and aDVM2). Two models always simulated regeneration for both the 7 and 10 cm threshold (xComp and TreeMig), i.e., they did not feature any zero values. The other ten models had a very low percentage of samples with no regeneration (0.01% - 2.39%), i.e., they had distinctly fewer occurrences of zeros compared to the observations.

@@

Most models matched the level of diversity of the observed data quite well (Figure \@ref(fig:H710)). Five models overestimated regeneration diversity: ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter particularly for the 7 cm diameter threshold. The model 4C is a special case, as it simulated five species only, i.e., its diversity values are not directly comparable to those of the other models, nor to the observations. Only one model, Landis-II, consistently underestimated regeneration diversity.

In most models, there were only small but significant differences in the species diversity of regeneration between 7 and 10 cm across sites (Tables S2). Four models (iLand, Landis-II, TreeMig and LPJ-GUESS) maintained the regeneration diversity between the 7 and 10 cm thresholds (i.e., the differences between them were not significant, see Table S2), and the same was evident from the observed data.

For both the observed and the simulated data, and both regeneration thresholds, species diversity in the regeneration compared to stand-level diversity did not feature clear patterns across the models (Figures S1 and S2). Most models captured reasonably well or overpredicted species diversity at the stand level (Figure \@ref(fig:combinedRichness)A, left and center; cf. Figure S3), but a characteristic feature was that the simulations had a much lower variability of diversity compared to observations. Models overpredicting species diversity included ForClim1, ForClim 11, FORMIND and TreeMig (Figure \@ref(fig:combinedRichness)A, left). Several models underpredicted species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure \@ref(fig:combinedRichness)A, right).

The majority of the models overpredicted regeneration diversity; as already observed in Figure \@ref(fig:H710), only two of them underestimated it (4C and Landis-II) (Figure \@ref(fig:combinedRichness)B and Figure S4). Both the models with feedback and those without feedback from the adult trees to regeneration (via seed production) overestimated, underestimated or captured reasonably well regeneration species diversity compared to the observations. Again, simulated diversity had considerably lower variability than observed diversity in the regeneration.

@@

There were strong differences among the models regarding the mortality rate between 7 and 10 cm (Figure \@ref(fig:mortality11)). Most of the models featured a mortality rate significantly larger than the observed data, thus at least partially compensating for the general overestimation of regeneration levels (Figure \@ref(fig:over) and Table S3); six models did not feature a significant difference compared to the observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2). Four models (xComp, PICUS, LandClim and TreeMig) featured very high mortality rates (i.e., well above the Reineke self-thinning line), which compensated for the strong initial overestimation of regeneration (cf. Figure \@ref(fig:over)). Interestingly, two models that underestimated overall regeneration levels (SIBYLA and Landis-II; cf. Figure \@ref(fig:over)) featured mortality rates that were close to but still above the self-thinning line (Figure \@ref(fig:mortality11)).

The fact that some models (and observed data) featured lower mortality than expected by Reineke's self-thinning rule may point to facilitation, or simply a higher mortality rate before the trees had reached 7 cm diameter. Yet, the case of models such as xComp, PICUS and TreeMig, whose mortality was well above the self-thinning line, indicates that they feature higher regeneration mortality compared to even-aged, single-species forests. Some models yielded a mortality rate of regeneration that is broadly compatible with the self-thinning rule (e.g., 4C and LandClim).

The three models with the largest overestimation of the proportion of regeneration at the 7 cm threshold featured the highest mortality between the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A). The other models that overestimated regeneration had mortalities lower than the expected self-thinning ratio, with the exception of LandClim (Figure \@ref(fig:mortality13)B). It is noteworthy that most of the models that featured a low ratio of regeneration between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small overestimation of regeneration at the 7 cm threshold (Figure \@ref(fig:mortality11)). aDGVM2 was the only model that underestimated regeneration at the 7 cm threshold while featuring a low mortality between the two thresholds (Figure \@ref(fig:mortality13)C), and only a few models were close to the correct regeneration levels at 7 cm while also being close to the Reineke line (SIBYLA, Landis-II; cf. Figure \@ref(fig:mortality13)D; and 4C as well as LandClim).

Lastly, some models that had a low number of individuals at the 7 cm threshold also had a smaller decrease of regeneration between 7 and 10 cm (ratio 7-10 cm) (Figure S5). In these models, competition and self-thinning are either not pronounced or must have occurred before the trees had reached 7 cm. However, this pattern was not consistent across models (Figure S5). It is noteworthy that there is a relationship between the regeneration levels at 7 cm and the mortality rate (Table S4): most of the models showed a positive effect (higher regeneration at 7 cm is coupled to an increase of the mortality between 7 and 10 cm). While some models showed a negative effect, the linear trend was not pronounced and not always significant regarding the slope.

@@

When evaluating regeneration levels along key gradients of light availability (basal area), temperature (degree-days) and soil moisture (climatic water balance), distinct features emerged: the models reproduced the effect of basal area in both its magnitude and patterns (Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better than the effects of the climatic gradients, where they featured varying patterns.

>>>>

Most of the models overestimate recruitment levels, which is only compensated in some models by high simulated mortality rates in the early stages of individual trees dynamics. Simulated species diversity of recruitment matches the observed ranges. Models simulating higher species diversity at the stand level do not present higher recruitment diversity. The effect of light availability on recruitment levels is captured better than the effect of temperature and soil moisture, but patterns are not consistent across models. Increasing complexity in the tree regeneration modules of the models is not related to higher accuracy of simulated tree recruitment. Furthermore, individual model design is more important than scale (stand, landscape, global) and approach (empirical, process-based) for accurately capturing tree regeneration.

@@

Thus, a focus on the modeling of tree regeneration processes is sorely needed and overdue [@price2001; @walck2011] if we are to make reliable projections of future forest dynamics, i.e., when the models need to be operated in extrapolation mode, as well as from a fundamental ecological point of view for increasing systems understanding. In the present study, for the first time, a large number of forest models commonly used to assess forest dynamics under climate change are evaluated against a continental-scale, multi-species harmonized dataset on tree recruitment [@käber2023]. By tree recruitment, we refer to the passing of trees across a specific diameter threshold ("ingrowth"). We included models that are based on a range of "philosophies", from models operating at the stand to the global scale as well as the range of models from empirically derived to "process-based" [e.g., @bugmann1996; @fabrika2012; @lexer2001; @reyer2014]. Due to the large variability in tree regeneration patterns in nature and the large number of factors driving and constraining this process --- including some that are not incorporated explicitly in most models, such as deer browsing --- we do not aim for a detailed statistical evaluation of each model. Instead, we aim to evaluate the general regeneration patterns and magnitudes simulated by the models, and benchmark the simulated regeneration niche of multiple species against empirical data along a wide environmental gradient of temperature, moisture, and light availability [@grubb1977].

More specifically, we aim to answer the following questions: (1) Are models of forest dynamics capturing accurately tree recruitment levels, initial tree species diversity and mortality in the recruitment? (2) Do model traits explain differences in model performance? (3) How well do the models capture total recruitment and individual species regeneration niches across environmental gradients of light availability, temperature, and soil moisture? We evaluate the performance of regeneration models by analyzing simulated data alongside observed data. Specifically, we examine tree species diversity, mortality rates in tree recruitment, and recruitment levels. Then, we link the model's traits and performance to define regeneration and we assess the recruitment gradients on regeneration niches. Finally, we provide research recommendations for next steps towards improving tree regeneration modules.

@@

In the models used here, the overall complexity in the regeneration modules varies considerably. Following @bugmann2022, we can classify seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and TreeMig) as having rather high complexity in their regeneration modules (mean regeneration formulation complexity across all processes \>0, Table \@ref(tab:tableModels)). Two models, ForClim and ForCEEPS, were used here with two alternative variants of regeneration. ForClim variant 1 [@bugmann1996] is based on a regeneration module that adheres closely to the concept introduced by @botkin1972 which is an environmental filter considering species specific thresholds of light availability and climatic variables. Whereas ForClim variant 11 is adopting a slightly more complex approach where individual species properties and their relation to the environment are represented gradually along with the relative suitability of different species (i.e., competition among tree species) [@huber2020]. These two model variants allow us to evaluate a more process-based and complex module (variant 11) against a simple module (variant 1), while the rest of the model structure is identical. Similarly, the two ForCEEPS variants allow us to isolate the importance of the canopy feedback (i.e., simulated actual composition and relative abundance of species in the plot) via seed trees for the quantity and quality (e.g., diversity and composition) of simulated regeneration, as one variant includes this feedback whereas the other does not.

@@

Recruitment data covering a wide range of environmental conditions are hard to obtain, and this is one of the reasons why most models of forest dynamics have never been confronted with a dataset covering such gradients over a large number of sites, to evaluate how well regeneration is captured. The observations used here are derived from a novel network of sites in forest reserves that represent the range of environmental gradients in temperature and precipitation in Central Europe as compiled in the framework of the EuFoRIa network [@euforia2019;@käber2023] (Figure \@ref(fig:map)). The environmental coverage and the variety of forests represented by this data is unprecedented within Europe. These forests have been unmanaged for at least 10 years prior to their designation as reserves, and most of them provide time series of natural forest dynamics over multiple decades and up to ca. 80 years. The census periods range from 3 to 37 years, with an average of 14 years. The data provide information at the tree level, thus allowing for the sequential comparison of processes such as individual-tree regeneration and death between the measurements.

We selected 200 sites from this network as the benchmarking dataset for the simulation to be representative of the environmental variation contained in the data. This was achieved by applying k-means clustering to define 200 clusters of plots from the original set of 869 plots along the environmental dimensions of temperature, climatic water balance, soil quality, slope and aspect. Each of these sites featured at least two consecutive measurements. Recruitment minimum size thresholds for these sites differed between stem diameters of 0 and 10 cm. For our study, we defined two datasets where one included 165 sites with a diameter threshold of 7 cm or lower, and the other included another 35 sites with diameter thresholds from 10 cm. Therefore we had in total 200 sites with observations of newly established trees above 10 cm and 165 sites above 7cm. In the results section, any result from observed data at 7 cm threshold is originated from the 165 sites.

Plot size ranged from 0.02 ha to 5.52 ha, and the data were further processed and aggregated following @käber2023 to provide recruitment rates per ha and per decade. The observations used in this study featured 30,900 newly established trees. Recruitment rates per site, sample and decade ranged from zero to 1246 trees, with a mean of 56 trees. Adult species composition was also available for each plot. For more details on this dataset and the detailed regeneration information, cf. @käber2023.

@@

The overarching goal of the simulation experiments was to assess tree recruitment as it arises from empirical data against its representation in a wide range of models of forest dynamics. We define tree recruitment as the passing of a breast height diameter threshold of 7 and 10 cm (synonym: ingrowth). To this end, each modeling group was provided with a detailed protocol [@díaz-yáñez2022] with instructions how to perform the simulations, input variables on climate and soil conditions, and the list of expected output variables. Neither were further site information (except for the data specified below) nor any data on tree recruitment or forest stand features provided prior to the simulation.

@@

The simulations were run in the absence of natural disturbances. Only the model LPJ-GUESS had to include a background disturbance to increase chances of shade-intolerant species to establish. The simulations were set up to sample species-specific recruitment rates per decade and per ha in the equilibrium state of the model, typically entailing a "spin-up" run (as we did not provide any forest data). The modeling teams decided on the simulated area and how they derived these samples. The exact length of the simulation was also decided by the modeling teams (Table \@ref(tab:tableModels)). Further details on how each modeling team prepared the simulations and the outputs are available in Supplementary Material 1.

The simulations were run in the absence of management to a simulated equilibrium ("Potential Natural Vegetation") with the current climate, as the ultimate goal was to evaluate tree recruitment under comparable and near-equilibrium conditions. This entails the assumption that (1) the observations from the forest reserves reflect no traces of forest management, and (2) there is an equilibrium between forest dynamics and climate. While the former might be starting to be visible in many of the EuFoRIA reserves, the latter may be more debatable. However, in the absence of detailed data on the history of each plot in the EuFoRIa network, some broad assumptions had to be made. Both the width of the regeneration niche (i.e., in environmental space) as well as the intensity of the regeneration process (i.e., the number of recruited trees per area and per unit of time) were of interest.

@@

Two models did not provide results from all the simulated samples or sites to avoid unrealistic results, as follows. In the model 4C, a threshold of a maximum basal area of 90 $m^2 ha^{-1}$ was used to avoid unrealistic stand basal area data, and therefore not all the samples and sites were reported. The reason for this is that 4C is not suitable for long-term simulations without management, due to misrepresentations in density-dependent mortality processes in long-term simulations and assumptions of tree geometry that lead to unrealistic single tree dimensions of very old individuals. LPJ-GUESS had 2% of the sample outputs without tree recruitment, and these were considered as zero stand basal area and zero regeneration for all the tree species simulated; one site produced grassland rather than a forest and was not included in the results.

@@

The simulation results were analyzed regarding (1) recruitment levels (i.e., ingrowth number per unit time and space), (2) recruitment species diversity, (3) recruitment mortality, (4) the relationship between model performance and model traits, and (5) ingrowth gradients along the regeneration niches. We evaluated species diversity across the models and in relation to the observed data by calculating the Shannon index $H$ based on the relative proportion of the species in terms of basal area. It was calculated for the regeneration ($H_{R_{n}}$) (Equation \ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$) (Equation \ref{eqpiShannon}). The higher the value of the index, the higher the species diversity at a particular site and sample.

@@

Mortality in tree recruitment was assessed based on the ratio of recruitment between the 7 and 10 cm diameter thresholds. We used the Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference to estimate whether the ratio of recruitment between the 7 and 10 cm diameter thresholds were above or below the expected theoretical rate. The Reineke self-thinning rule is usually calculated for even-aged, single species stands and assumes a fixed relationship between the number of stems and the quadratic mean diameter in fully stocked pure stands. The value used in our comparisons was 1.77 (i.e., we expect stem numbers at 7 cm to be 77% higher than at 10 cm), calculated using Equation \ref{eqReineke}.

```{=tex}
\begin{equation}
Reineke = \frac{N7}{N10}^{-1.605} \label{eqReineke}
\end{equation}
```
We assessed model performance in relation to model traits focusing on (1) model complexity as defined by @bugmann2022, (2) model type (empirical or process based), (3) the presence or absence of a canopy feedback for regeneration, and (4) the scale of application of the model (stand, landscape or global). We tested for significant differences between simulated and observed values of the recruitment levels and species diversity by using paired t-tests.

Total recruitment and the regeneration niches of the individual species were evaluated across the environmental gradients of light availability, temperature, and soil moisture. Stand basal area (the basal area of all the trees in each sample including tree regeneration) was used as a proxy for light availability at the forest floor, the annual degree-day sum [@allen1976; @fischlin1995] as a proxy for growing season warmth, and the climatic water balance as a proxy for soil moisture [@speich2019]. Recruitment values of each species were calculated as the mean across the 200 samples per site. The observed data were modeled using a Generalized Additive Model [@wood2011] with a negative binomial distribution and restricted maximum likelihood to better understand the relationship between the environmental gradients and the levels of observed regeneration, relative to the simulation results per model.

In order to analyze the regeneration niches across the climatic gradients, we focused on five common tree species or genera: *Abies alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and *Quercus* spp. For these species, we calculated the share in the recruited basal area per site, where each site represents a part of the environmental space, as the mean across the available samples per site:

```{=tex}
\begin{equation}
\bar{R}_{BA_{share_{t,i}}} = \frac{\bar{r}_{BA_{t,i}}}{\sum_{i = 1}^{s}{\bar{r}_{{BA_{t,i}}}}} \label{eq:BAshare}
\end{equation}
```
where $s$ is the total number of species simulated, and $\bar{r}_{t,i}$ is the mean basal area in the recruitment (subsequently referred to as "recruited basal area") of species $i$ at site $t$ across the available samples at that site. Furthermore, regarding the share in the recruited  basal area per species ($\bar{R}_{BA_{share_{i}}}$), we categorized this as zero when the recruited basal area of that species ($\bar{r}_{BA_{t,i}}$) was zero, but also when both the total recruited basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$) and the recruited basal area of that species ($\bar{r}_{BA_{t,i}}$) equaled zero.

@@

Simulated levels of recruitment varied strongly across the 15 models and typically did not match the levels found in the forest reserve data (Figure \@ref(fig:over)). Recruitment was overestimated in most models for both the 7 and 10 cm diameter thresholds, with the exception of the empirical stand model SIBYLA, the landscape model Landis-II and the global model aDGVM2, which estimated recruited levels at the lower end of the plausibility interval of the observed data (Figure \@ref(fig:over)). The models with the largest overestimation were the stand model PICUS and the landscape model TreeMig. For most models, the variability of simulated regeneration levels across the 200 sites (visible from the interquartile range in the box plots of Figure \@ref(fig:over)) was similar to or smaller than observed, with the notable exception of PICUS, where simulated recruitment variability was much larger.

Both the observations and the simulated data had no recruitment in some samples and at some sites (for details, cf. Table S1). The observed data had 4% of the samples with no recruitment. Only three models had a larger proportion of no recruitment (4C, Landis-II and aDVM2). Two models always simulated recruitment for both the 7 and 10 cm threshold (xComp and TreeMig), i.e., they did not feature any zero values. The other ten models had a very low percentage of samples with no recruitment (0.01% - 2.39%), i.e., they had distinctly fewer occurrences of zeros compared to the observations.

@@

Most models matched the level of diversity of the observed data quite well (Figure \@ref(fig:H710)). Five models overestimated recruitment diversity: ForCEEPS, ForClim 1, PICUS, TreeMig and LandClim, the latter particularly for the 7 cm diameter threshold. The model 4C is a special case, as it simulated five species only, i.e., its diversity values are not directly comparable to those of the other models, nor to the observations. Only one model, Landis-II, consistently underestimated recruited diversity.

In most models, there were only small but significant differences in the species diversity of recruitment between 7 and 10 cm across sites (Tables S2). Four models (iLand, Landis-II, TreeMig and LPJ-GUESS) maintained the recruitment diversity between the 7 and 10 cm thresholds (i.e., the differences between them were not significant, see Table S2), and the same was evident from the observed data.

For both the observed and the simulated data, and both recruitment thresholds, species diversity in the recruitment compared to stand-level diversity did not feature clear patterns across the models (Figures S1 and S2). Most models captured reasonably well or overpredicted species diversity at the stand level (Figure \@ref(fig:combinedRichness)A, left and center; cf. Figure S3), but a characteristic feature was that the simulations had a much lower variability of diversity compared to observations. Models overpredicting species diversity included ForClim1, ForClim 11, FORMIND and TreeMig (Figure \@ref(fig:combinedRichness)A, left). Several models underpredicted species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure \@ref(fig:combinedRichness)A, right).

The majority of the models overpredicted recruitment diversity; as already observed in Figure \@ref(fig:H710), only two of them underestimated it (4C and Landis-II) (Figure \@ref(fig:combinedRichness)B and Figure S4). Both the models with feedback and those without feedback from the adult trees to regeneration (via seed production) overestimated, underestimated or captured reasonably well recruitment species diversity compared to the observations. Again, simulated diversity had considerably lower variability than observed diversity in the recruitment.

@@

There were strong differences among the models regarding the mortality rate between 7 and 10 cm (Figure \@ref(fig:mortality11)). Most of the models featured a mortality rate significantly larger than the observed data, thus at least partially compensating for the general overestimation of recruitment levels (Figure \@ref(fig:over) and Table S3); six models did not feature a significant difference compared to the observed data (ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2). Four models (xComp, PICUS, LandClim and TreeMig) featured very high mortality rates (i.e., well above the Reineke self-thinning line), which compensated for the strong initial overestimation of recruitment (cf. Figure \@ref(fig:over)). Interestingly, two models that underestimated overall recruitment levels (SIBYLA and Landis-II; cf. Figure \@ref(fig:over)) featured mortality rates that were close to but still above the self-thinning line (Figure \@ref(fig:mortality11)).

The fact that some models (and observed data) featured lower mortality than expected by Reineke's self-thinning rule may point to facilitation, or simply a higher mortality rate before the trees had reached 7 cm diameter. Yet, the case of models such as xComp, PICUS and TreeMig, whose mortality was well above the self-thinning line, indicates that they feature higher recruitment mortality compared to even-aged, single-species forests. Some models yielded a mortality rate of recruitment that is broadly compatible with the self-thinning rule (e.g., 4C and LandClim).

The three models with the largest overestimation of the proportion of recruitment at the 7 cm threshold featured the highest mortality between the 7 and 10 cm thresholds (Figure \@ref(fig:mortality13)A). The other models that overestimated recruitment had mortalities lower than the expected self-thinning ratio, with the exception of LandClim (Figure \@ref(fig:mortality13)B). It is noteworthy that most of the models that featured a low ratio of regeneration between 7 and 10 cm (i.e., ForCEEPS, ForClim 1, ForClim 11 and iLand) had only a small overestimation of recruitment at the 7 cm threshold (Figure \@ref(fig:mortality11)). aDGVM2 was the only model that underestimated recruitment at the 7 cm threshold while featuring a low mortality between the two thresholds (Figure \@ref(fig:mortality13)C), and only a few models were close to the correct recruitment levels at 7 cm while also being close to the Reineke line (SIBYLA, Landis-II; cf. Figure \@ref(fig:mortality13)D; and 4C as well as LandClim).

Lastly, some models that had a low number of individuals at the 7 cm threshold also had a smaller decrease of recruitment between 7 and 10 cm (ratio 7-10 cm) (Figure S5). In these models, competition and self-thinning are either not pronounced or must have occurred before the trees had reached 7 cm. However, this pattern was not consistent across models (Figure S5). It is noteworthy that there is a relationship between the recruitment levels at 7 cm and the mortality rate (Table S4): most of the models showed a positive effect (higher regeneration at 7 cm is coupled to an increase of the mortality between 7 and 10 cm). While some models showed a negative effect, the linear trend was not pronounced and not always significant regarding the slope.

@@

When evaluating recruited levels along key gradients of light availability (basal area), temperature (degree-days) and soil moisture (climatic water balance), distinct features emerged: the models reproduced the effect of basal area in both its magnitude and patterns (Figure \@ref(fig:nvTrend710BAlightTemp) and S7) considerably better than the effects of the climatic gradients, where they featured varying patterns.

In the observed data, recruited levels decreased clearly with increasing stand basal area (Figure \@ref(fig:nvTrend710BAlightTemp)a). This trend was captured by the Landis-II model only, albeit at much lower values of basal area than in the observations. The other models featured distinctly different trends, such as (1) an increase in recruited levels with increasing total basal area followed by the absence of regeneration at high values of basal area (ForCEEPS, ForCEEPS(f), FORMIND, TreeMig and LPJ-GUESS), (2) almost constant recruitment levels with basal area (ForClim 11, xComp, LandClim and aDGVM2), or (3) an increase of recruitment up to a certain value of stand basal area followed by a decrease at even higher values of stand basal area, with model-specific thresholds (4C, ForClim 1, PICUS, SIBYLA and iLand).
